---
title: "Week 10 Live Session Assignment - Website Click Data"
author: "Ramesh Simhambhatla"
date: "November 13, 2016"
output: 
  html_document:
    keep_md: true
---

# Introduction
The purpose of this report to is to produce code and output for Week10 Live Session Assignment on the WebSite Click Data. The data set used in nyt1.csv

####Load libraries used in this exercise
```{r}
library(plyr)
library(dplyr)
library(data.table)
library(ggplot2) # used for visualizations
```
####The read the nyt1.csv from the columbia.edu site and store into a local data frame 'data1'
```{r}
fileLocation <- "http://stat.columbia.edu/~rachel/datasets/nyt1.csv"
data1 <- read.csv(url(fileLocation))
```

####Explore the data using head, str, and summary functions
```{r}
head(data1) # reads top6 observations
str(data1) # str stands for 'structure'
summary(data1) # summary of the variables

# distribution of the Age column
hist(data1$Age, main="", xlab="Age")
range(data1$Age)

# distribution of the Impressions column
hist(data1$Impressions, main="Distribution of the Impressions", xlab="# of Impressions")

range(data1$Impressions)

# distribution of the Clicks column
hist(data1$Clicks, main="", xlab="# of Clicks")

range(data1$Clicks)
```

####Create a new variable ageGroup that categorizes age into following groups: < 18, 18???24, 25???34, 35???44, 45???54, 55???64 and 65+.

```{r}
data1$Age_Group <- cut(data1$Age, c(-Inf, 18, 24, 34, 44, 54, 64, Inf))
levels(data1$Age_Group) <- c("<18", "18-24", "25-34", "35-44", "45-54", "55-64", "65+")
# Name the levels of 'Age_Group' for readability
head(data1)
```

####Use sub set of data called ???ImpSub??? where Impressions > 0 ) in the data set.
```{r}
ImpSub <- subset(data1, Impressions>0) # create a subset for Impressions > 0
ImpSub$CTR <- ImpSub$Clicks/ImpSub$Impressions
head(ImpSub)
```

####Create a new variable called click-through-rate (CTR = click/impression).
####Use this ImpSub data set to do further analysis.
```{r}
# Define a new variable to segment users based on click -through- rate (CTR) behavior.
# CTR< 0.2, 0.2<=CTR <0.4, 0.4<= CTR<0.6, 0.6<=CTR<0.8, CTR>0.8
ImpSub$CTRSeg <- cut(ImpSub$CTR, c(-Inf, 0.2, 0.4, 0.6, 0.8, Inf))
levels(ImpSub$CTRSeg) <- c("0.0-0.2", "0.2-0.4", "0.4-0.6", "0.6-0.8", "0.8-0.1")
head(ImpSub)
```

####Get the total number of Male, Impressions, Clicks and Signed_In (0=Female, 1=Male)
```{r}
ImpSubMale <- subset(ImpSub, ImpSub$Gender==1) # create a subset for Gender = 1(Male)
sapply(ImpSubMale[, 2:5], sum) # output sum for columns 2 through 5
```

####Get the mean of Age, Impressions, Clicks, CTR and percentage of males and signed_In
```{r}
sapply(ImpSub[,c(1,3,4,7)], mean) # mean of Age, Impressions, Clicks, CTR
pcntMales <- length(ImpSubMale$Gender)/length(ImpSub$Gender) # calculate percentage of Males
pcntMales # print percentage of Males

ImpSubSignedIn <- subset(ImpSub, ImpSub$Signed_In==1) # create subset for signed_in = 1 data
pcntSignedIn <- length(ImpSubSignedIn$Signed_In)/length(ImpSub$Signed_In) # calculate the percent signed_in
pcntSignedIn # output percent Signed In

#summarise to print consolidted data for mean and percentages 
summarise(ImpSub, mean(Age), mean(Impressions), mean(Clicks), mean(CTR), pcntMales, pcntSignedIn)
```
####Get the means of Impressions, Clicks, CTR and percentage of males and signed_In  by AgeGroup.
```{r}
dataVByAgeGroup <- ImpSub %>% 
  group_by(Age_Group) %>% 
  summarise(Impressions=mean(Impressions), CTR=mean(CTR), Clicks=mean(Clicks), 
            pctSigned=sum(Signed_In==1)/length(Age_Group),
            pctMales=sum(Gender==1)/length(Age_Group))

dataVByAgeGroup
```
####Create a table of CTRGroup vs AgeGroup counts.
```{r}
countAgeGroup <- as.data.table(ddply(ImpSub, .(Age_Group), nrow))
names(countAgeGroup)[2] <- "countAgeGroup"
countAgeGroup

countCTRGroup <- as.data.table(ddply(ImpSub, .(CTRSeg), nrow))
names(countCTRGroup)[2] <- "countCTRGroup"
countCTRGroup

mergeCounts <- rbind(countCTRGroup, countAgeGroup, fill=TRUE)
mergeCounts
```

#### Plot distributions of number impressions and click-through-rate (CTR = click/impression) for the age groups
```{r}
ggplot(subset(ImpSub, Impressions>0), aes(x=Impressions, fill=Age_Group)) +
  geom_histogram(binwidth=1)

ggplot(subset(ImpSub, CTR>0), aes(x=CTR, fill=Age_Group)) +
  labs(title="Click-through rate by age group - for a Single Day") +
  geom_histogram(binwidth=.025)
  
# Histogram of Clicks by Age
ggplot(subset(ImpSub, Age>0), aes(x=Age, fill=Age_Group)) +
  labs(title="Clicks by Age - for a Single Day") +
  geom_histogram(binwidth=0.5)
```
####**** End of File ****